outline

### 2019-11-21 Rmd/jse-2019-ms.Rmd
updated ms

---
title: 'A framework for mentored data science research projects'
author: "Frederick J. Boehm and Bret M. Hanlon"
date: "8/20/2019"
params:
  lastmod: !r lubridate::now()
bibliography: jse.bib
output: bookdown::html_document2
---

>>>>

---
title: 'A framework for mentored data science research'
author: "Frederick J. Boehm and Bret M. Hanlon"
date: "8/20/2019"
params:
  lastmod: !r lubridate::now()
bibliography: jse.bib
output: 
  - github_document
  - pdf_document
---

### 2019-11-21 Rmd/methods.Rmd
recompiled pdf and md
updated ms

@@

Our computational system for acquiring tweets involved several steps. We interacted 
with the API via the R package `twitteR` [@twitteR]. We used the free Twitter streaming 
API that gave us access to approximately one percent stream of all tweets during the 
specified query time period. To ensure that we collected tweets continuously, we used 
the linux tool `crontab` to execute our R script every five minutes. Each execution of 
the R script performed a single streaming API query, lasting five minutes. Twitter's 
streaming API, at the time of our data collection, enforced rate limits on the 
frequency and duration of queries. With the above settings, we continuously collected 
tweets.

@@

Our framework, and the projects developed under it, encourage students to deepen 
computational reasoning skills in several ways. First, they work with a variety of 
internet-based data to answer research questions. In the two example cases, our 
students collected tweets over time and gathered complementary data from other 
resources, including daily closing prices of stock market indexes. This gave students 
opportunities to think creatively about what data to acquire and how to use multiple 
data sources in a single cohesive project.

Second, the students worked with a variety of data structures. 
The Twitter streaming 
API returns tweets as JSON (Javascript Object Notation, http://json.org). Because 
distinct Twitter users may provide 
different pieces of profile information, there is variability in the structure of 
each tweet's JSON. Additionally, tweet metadata fields may appear in any order (https://developer.twitter.com/en/docs/tutorials/consuming-streaming-data). 
Students needed to recognize this and write code that 
accommodated these variations in tweet data structure.

@@

**see https://developer.twitter.com/en/docs/tweets/data-dictionary/overview/tweet-object**

>>>>

Additionally, our projects gave students opportunities to develop and to practice skills in reproducible research. Given the growing imperative to document and share code, we feel that this skill set equals in importance the three points above.

4. develop and practice skills in reproducible research

Below, we describe how our framework enabled students to achieve in the four items listed above.

@@

Our computational system for acquiring tweets involved several steps. We interacted 
with the API via the R package `twitteR` [@twitteR]. We used the free Twitter streaming 
API that gave us access to approximately one percent stream of all tweets during the 
specified query time period. To ensure that we collected tweets continuously, we used 
the linux tool `crontab` to execute our R script every five minutes. Each execution of 
the R script performed a single streaming API query for five minutes. Twitter's 
streaming API, at the time of our data collection, enforced rate limits on the 
frequency and duration of queries. With the above settings, we continuously collected 
tweets.

@@

Our framework encourages students to deepen 
computational reasoning skills in several ways. First, they work with a variety of 
internet-based data to answer research questions. In the two example cases, our 
students collected tweets over time and gathered complementary data from other 
resources, including daily closing prices of stock market indexes. This gave students 
opportunities to think creatively about what data to acquire and how to use multiple 
data sources in a single cohesive project.

Second, the students worked with a variety of data structures. 
The Twitter streaming 
API returns tweets as JSON (Javascript Object Notation, http://json.org). Because 
distinct Twitter users may provide 
different pieces of profile information, there is variability in the structure of 
each tweet's JSON. 
Additionally, tweet metadata fields may appear in any order (https://developer.twitter.com/en/docs/tutorials/consuming-streaming-data). 
Students needed to recognize this and to write code that 
accommodated these variations in tweet data structure.

@@

**see https://developer.twitter.com/en/docs/tweets/data-dictionary/overview/tweet-object**
**insert a 2-column table of fields here**
